Digital video platforms and phylogenetic modelling reveal parrot tool use is not rare

### 2024-07-12 manuscript.Rmd
revise and resubmit

@@

Comparative work on the evolution of tool use is a paradigmatic example of this 
issue. The initial discoveries of tool use in chimpanzees [@Goodall1964], 
birds [@Hunt1996], dolphins [@Smolker1997], and octopuses [@Finn2009] occurred 
decades after significant advances on other more easily measurable aspects of 
their biology. Since then, scholars have proposed a clear operational 
definition of tool use applicable to all species --- the manipulation of an 
unattached object as an extension of the animal's body to achieve a 
goal [@Shumaker2011] --- and have used the distribution of species meeting this 
definition to make various claims about the evolutionary drivers of tool use 
behaviours. For example, based on the observation that bird species with 
reported tool use tended to have larger brains, researchers have identified 
higher relative brain size as a likely precondition for tool using 
capabilities [@Lefebvre2002;@Lefebvre1997;@Lefebvre2004]. These researchers 
argue that larger brains are better able to integrate visual and somatosensory 
information when innovating novel behaviours, such as tool use, in changing 
environments [@Sol2005;@Sol2009]. Similarly, researchers have used existing 
reports of tool use in birds to debate the roles of generalist versus 
specialist feeding strategies in driving the evolution of tool use, with some 
arguing that feeding generalists require technical innovations to expand their 
dietary niche [@Ducatez2015;@Overington2011;@Sol2005] and others arguing that 
feeding specialists require technical innovations for extractive foraging of 
specific foods [@HenkeVonDerMalsburg2020;@MettkeHofmann2014].

However, before we can make claims of this kind, we need to know whether current 
research effort in the literature is sufficient for robust conclusions to be 
drawn about the evolution of tool use. In fact, evidence suggests that research 
effort is often systematically biased towards particular taxonomic groups, parts 
of the world that are easy to access, and species with life history traits that 
make them easier to study, such as larger distribution ranges and population 
sizes [@Ducatez2014]. This is a crucial limitation because insufficient 
observation may lead researchers to miss true instances of tool behaviours and 
thus draw premature conclusions about the evolutionary drivers and origins of 
tool use. While researchers have attempted to deal with this problem by 
controlling for the number of scientific papers published on different species, 
previous work has not yet attempted to quantify and explicitly model the 
relationship between actual tool-using behaviour and what is reported in the 
scientific literature. If more tool-using species exist than previously thought, 
this could have important implications for theories of the evolutionary drivers 
and origins of tool use and for our understanding of how rare this behaviour 
actually is.

@@

```{r echo=F}
tar_load(crowdData)
trueCrowdData <- filter(crowdData, ToolUse == "Y")
bordCrowdData <- filter(crowdData, ToolUse == "B")
```

@@

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common.

@@

In our model, we infer the tool-using status of each species by allowing each 
species to have their own probability of being a non-tool-user. Following our 
causal model (Figure \@ref(fig:plotDAG)), we predict these probabilities based 
on feeding strategy, encephalisation quotient, and phylogenetic history (see 
Methods for full model). The model additionally takes research effort into 
account by allowing that, among species for which tool use is unobserved, all 
else being equal those with fewer published papers and fewer video search hits 
have a higher probability of being undetected tool users (Supplementary Figure 
\@ref(fig:plotSurvCure3)).

@@

We found that this phylogenetic survival cure model was able to adequately 
distinguish between species with and without evidence for tool use, with an 
area-under-the-curve classification statistic of 
`r printnum(tar_read(auc), digits = 2)` (Supplementary Figure 
\@ref(fig:plotSurvCure9)). To further estimate the accuracy of the model's 
predictions, we also used a leave-one-species-out approach with known tool 
users. For each of the 25 tool-using species that were represented on the 
phylogeny and for which we had brain size and genomic data (we lacked data for 
three tool-using species), we fitted the model to a modified dataset which set 
tool use to be absent for the target species in both the scientific literature 
and the video survey. Across `r length(tar_read(cv))` cross-validation models, 
`r sum(tar_read(cv))` models (`r printnum(mean(tar_read(cv))*100, digits = 0)`%) 
continued to predict the target species as having a median posterior probability 
of tool use that was within the range of all other tool users. This 
classification rate was greater than the baseline classification rate of 
`r printnum(mean(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274) * 100, digits = 0)`% 
for species without evidence of tool use in the full model
(`r sum(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274)` 
of `r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
evidence of tool use had a median posterior probability of tool use that was 
within the range of the tool-using species). Together, the area-under-the-curve 
statistic and the leave-one-species-out approach suggest that the model is able 
to adequately classify known tool users, with some error.

Figure \@ref(fig:plotSurvCure1) visualises the ranked posterior probabilities of 
tool use from the phylogenetic survival cure model for all parrot species. As 
expected, the known tool users are ranked towards the top of this list. However, 
several "tool use absent" species also rank highly on the list, despite not 
being identified as tool users in the scientific literature or in our video 
survey. In fact, according to the model, the most likely tool user is a species 
for which tool use is unobserved in our data: the blue-eyed cockatoo (_Cacatua 
opthalmica_). This species is endemic to Papua New Guinea and is relatively 
understudied, with only `r tar_read(d)$EX.Pre.Tool.Papers[165]` published papers 
and `r tar_read(d)$SearchHits[165]` video search hits, which is fewer than the 
model expects are necessary to discover tool use when it is present (Figure 
\@ref(fig:plotSurvCure2)). This species is also found in the _Cacatua_ genus, a 
clade containing several known tool users. This prediction makes sense given the 
high phylogenetic signal for tool use reported by the model (Supplementary 
Figures \@ref(fig:plotSurvCure5) and \@ref(fig:plotSurvCure6)). Beyond the 
blue-eyed cockatoo, other highly ranked species without observed evidence of 
tool use are the Meyer's parrot (_Poicephalus meyeri_), the golden parakeet 
(_Guaruba guarouba_), the long-billed corella (_Cacatua tenuirostris_), the 
Solomons cockatoo (_Cacatua ducorpsii_), the red-fronted parrot (_Poicephalus 
gulielmi_), the Cape parrot (_Poicephalus robustus_), the yellow-eared parrot 
(_Ognorhynchus icterotis_), the red-vented cockatoo (_Cacatua haematuropygia_), 
and the gang-gang cockatoo (_Callocephalon fimbriatum_). Figure 
\@ref(fig:plotPhylo1) plots these species on the parrot phylogeny, using the top 
ten highest ranked species without observed evidence of tool use as an arbitrary 
cutoff for visualisation purposes.

@@

The predicted probabilities from our phylogenetic survival cure model have 
implications for inferences about the evolutionary drivers and origins of tool 
use in the parrot order. Regarding the drivers of tool use hypothesised in 
Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that 
encephalisation quotient strongly positively predicted the probability of tool 
use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% 
CI [`r printnum(quantile(-postFull$bEQ, 0.025))` 
`r printnum(quantile(-postFull$bEQ, 0.975))`]; total _n_ = `r nrow(tar_read(d))` 
species; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in 
Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative 
brain size in the dataset. We also found that feeding generalist species were 
slightly more likely to be tool users, though the posterior difference between 
generalists and specialists was quite uncertain (median posterior log odds 
difference = `r printnum(median(diff))`, 95% CI [
`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]; total 
_n_ = `r nrow(tar_read(d))` species). These results from the survival cure model 
differed from the results of models fitted to pre-video-survey and 
post-video-survey data without the survival cure component, which found no 
effect of relative brain size and no difference between feeding strategies, 
respectively (Supplementary Figure \@ref(fig:plotComparison)).

@@

Regarding the origins of tool use, we fitted exploratory ancestral state 
reconstruction models to the pre-video-survey data, the post-video-survey data, 
and the predicted probabilities from the phylogenetic survival cure model. The 
discoveries from our video survey and from our phylogenetic modelling increased 
the likelihood that tool use was present in the most recent common ancestor for 
several parrot genera. These include the amazon parrots native to the Americas 
(_Amazona_), the true white cockatoos and corellas found in South East Asia and 
Australasia (_Cacatua_), the kea and the kākā from New Zealand (_Nestor_), and 
the _Poicephalus_ genus native to Africa (Table \@ref(tab:tableASR); 
Supplementary Figures \@ref(fig:plotASR1) - \@ref(fig:plotASR3)). These findings 
suggest that species from each of these genera may share their tool use 
capabilities via common descent from their respective common ancestors, rather 
than via independent evolution within each genus or behavioural innovation
within species.

@@

These findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified. This raises the 
possibility that insufficient research effort is a general issue across the 
scientific literature, both for tool use in other groups and for other rare 
behaviours.

Second, in terms of the evolution of tool use in parrots, our study challenges a 
key assumption made in the literature to date: that only a minority of parrots 
are tool users [@Auersperg2012;@Bastos2021;@Goodman2018;@Lambert2015;@OHara2021]. 
The paucity of evidence for tool use across parrots in the literature initially 
implied that the capacity for tool use may have evolved independently in 
different parrot species. Our discovery of the widespread distribution of tool 
use across the parrot phylogeny, along with the strong phylogenetic signal in 
this expanded dataset, challenges this and suggests that, at least for some 
parrot clades, the capacity for tool use might be a homologous trait that has 
been evolutionarily conserved. Our exploratory ancestral state reconstruction 
analysis provides preliminary support for this hypothesis, revealing an 
increased probability of tool use among the most recent common ancestors for the 
_Amazona_, _Cacatua_, _Nestor_, and _Poicephalus_ genera. Even at this 
preliminary stage, our analysis therefore raises an alternative hypothesis for 
the observed tool use in _Cacatua_ [@Auersperg2012;@OHara2021] and 
_Nestor_ [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Bastos2021;@Goodman2018], namely that tool behaviours have arisen due to the 
common ancestor having the capacity to use tools, rather than from independent 
evolution or behavioural innovation within species.

@@

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and preening [@Goodman2018;@OHara2021], parrots in captivity
only require tools to achieve the latter goal. Despite this difference in the
type of tool use observed, self-scratching is nonetheless a major category of 
animal tool use. Self-scratching also meets the definition for a more complex 
type of embodied tool use known as 'tooling', the deliberate generation of a 
mechanical interface by using an object to manipulate another target or surface
[@Fragaszy2018]. This makes it difficult to dismiss self-scratching as 
unimportant: to do so would require a redefinition of tool-use in non-human
animals. Moreover, the fact that we found strong phylogenetic clustering of 
self-scratching tool use from the video survey with other examples of tool use
from the literature supports a common underlying cognitive mechanism. In line 
with this, some of the species of parrots that use tools for self-scratching in
captivity also use tools for other purposes in the wild [@Auersperg2012;
@Bastos2021;@Goodman2018;@OHara2021].

There was little evidence to suggest that the observations of self-scratching
tool use in the video survey were merely unintentional accidents or explicitly 
trained behaviours. Individual parrots often used tools slowly and repetitively 
over long periods of time, even across multiple different videos, suggesting 
that their behaviour was not random or accidental [@Sandor2020]. Parrots 
preferentially employed self-scratching tools on areas of their body that were 
otherwise inaccessible, with 96% of all instances involving scratching of the 
head or neck, suggesting intentional tool use. For 60% of the species in the 
video survey, we found two or more videos of repetitive and sustained scratching 
by different individuals of the same species in separate households, suggesting 
that the manipulations were intentional and recurring events that did not 
represent unusual stereotypies of any single individual. Regarding the 
possibility of training or cueing from owners, over half of the videos contained 
no evidence of human interaction aside from filming the behaviour. Humans only 
handed parrots their tools in two of the videos, and none of the videos featured 
owners directly rewarding tool use behaviours with food. It is unlikely that 
owners would specifically train parrots to scratch themselves with tools. 
Finally, the high levels of phylogenetic signal in our data provide strong 
evidence that the observations from our video survey reflect
biologically-endowed capacities for tool use rather than accidental or trained
behaviours, which would likely appear uniformly across the phylogeny.

@@

# Methods

## Video searches and coding

Our video search was conducted on YouTube in July 2020. Searches were conducted
manually by the first author over a month-long period, using the same IP address
and not logged in with a YouTube user account. Search terms included 
"parrot using tool" and variants (e.g., "macaw using tool", "lorikeet using 
tool", "parakeet using tool"), "tool use in parrot", "parrot tool use", "parrot 
scratching itself" (included after we found several videos demonstrating 
self-care tool use in previous searches) and equivalent terms (e.g., "parrot 
preening itself", "parrot grooming itself", "parrot scratching"). For all 
species that did not display results including object manipulation or scratching 
behaviours, we also searched the species’ common name(s) + "tool use", as well 
as the species’ common name(s) + "scratching". We also searched for translations 
of the terms "parrot tool use" and "parrot scratching" in languages for all 
countries where bird ownership was reported as >5% [@GlobalGfkSurvey], namely, 
Turkish, Czech, Polish, French, Italian, Dutch, German, Russian, Spanish, 
Portuguese, and Mandarin. Browser search histories were not cleared between
searches.

When we found a relevant video, we also searched for similar content uploaded by 
the same person/channel. For each YouTube search conducted, we watched all 
relevant videos until we reached five consecutive videos that did not feature 
any parrots. At this point, we ended that search and initiated the next search. 
In line with previous recommendations [@Nelson2013], we planned to exclude any 
videos that consisted of four or more shots edited together so as to ensure the 
behaviours being observed were not edited or manipulated, but none of the videos 
obtained qualified for exclusion.

All videos featuring parrots manipulating objects were investigated for 
potential tool use or borderline tool use. We defined tool use as the 
manipulation of an unattached object as an extension of the beak or foot to 
achieve a goal towards another object, individual, or oneself [@Shumaker2011]. 
Borderline tool use was similarly defined, except that it involved the use of an 
object that was still attached to a substrate [@Seed2010]. For example, if 
individuals used a fallen feather or stick for self-scratching this was defined 
as tool use, but using one’s currently attached tail feathers or cage 
furnishings for the same purpose was defined as borderline tool use. 
Self-scratching had to involve slow and repeated movements of touching an object 
to one’s body (or, in the case of borderline tool use, rubbing repetitively 
against an attached object [@Sandor2020]).

All relevant videos were coded for video length, species, tool use presence 
(yes/borderline), tool use type (e.g., scratching, feeding), the object being 
used (e.g., feather, stick), tool use target, human interaction 
(talking or handing object to parrot, holding parrot), and the number of shots 
within each video. Our complete dataset also includes the name for each video, 
link, subject name, sex (as declared by owner, as most parrot species are not 
sexually dimorphic), publishing date, and dates found and coded.

Since YouTube is constantly growing and evolving in both its uploaded content 
and its proprietary recommendation algorithms, the same search strategy today 
would not return exactly the same videos as our search in July 2020. It is 
possible that some videos have since been deleted or more videos of parrot tool
use have since been uploaded. However, this does not affect our conclusions
about the number of parrot species found to use tools in our initial search.

## Data for parrot species

We collected data for 194 parrot species (Supplementary Figures
\@ref(fig:plotPhylo2) - \@ref(fig:plotPhylo4)). We gathered feeding strategy 
data as a dichotomous variable ("generalist" or "specialist") from the 
EltonTraits ecological database [@Wilman2016]. As per the database, specialists 
were defined as species whose diet comprised at least 70% of a single food 
source. To calculate relative brain size, we collated data from the literature 
for all known body mass (g) and brain mass (g) values across 
parrots [@Flammer2001;@Iwaniuk2005;@Mazengenya2018;@Olkowicz2016;@Schuck2008;
@Silva2017]. For all species for which we obtained body and brain mass data, we 
calculated the encephalisation quotient (EQ) using the following 
formula [@Jerison1973]: $BrainWeight / (0.12 * BodyWeight(\frac{2}{3}))$. We 
found body mass and brain mass data for a total of 194 parrot species. This 
included all tool-using species in our video dataset, with the exception of 
three species: _Diopsittaca nobilis_, _Psittacara erythrogenys_, and _Coracopsis 
vasa_. For the latter, we used values for the closely related _Coracopsis 
nigra_. The other two species were excluded from the final dataset.

For modelling purposes, we coded research effort in both the scientific 
literature and the YouTube videos. For the scientific literature, we 
operationalised research effort as the number of papers published for each 
species' Latin name up to and including the first paper containing tool use for 
that species. If no tool use had been identified in the scientific literature 
for a species, then we coded the total number of papers published to date. We 
used the scientific database Scopus for coding the number of published papers. 
For the videos, we coded research effort as the number of search hits for each
species on YouTube. If tool use had been identified on YouTube, we estimated the
number of search hits when the first video of tool use was published on YouTube,
assuming linear growth of search hits since the inception of YouTube. If tool
use had not been identified, we used the current number of search hits.

For phylogenetic data, we used the phylogenetic tool at www.birdtree.org
[@Jetz2012] to compile 1000 posterior draws of phylogenetic trees for 174 of the 
194 parrot species for which both EQ and genomic data exist. A single maximum 
clade credibility tree was generated from these posterior draws for 
visualisation purposes. In our analyses, we iterated over posterior draws of the 
phylogeny to account for phylogenetic uncertainty.

## Phylogenetic signal

We used the _fitDiscrete_ function in the _ape_ R package [@Paradis2019] to 
calculate phylogenetic signal, for both the pre-survey and post-survey tool use 
data. We iterated the model over 100 posterior parrot phylogenies to incorporate 
phylogenetic uncertainty.

## Causal model of tool use

>>>>

\newpage
\linenumbers

```{r include=FALSE, message=FALSE, warning=FALSE}
library(papaja)
library(targets)
library(tidyverse)
```

@@

Comparative work on the evolution of tool use is a paradigmatic example of this 
issue. The initial discoveries of tool use in chimpanzees [@Goodall1964], 
birds [@Hunt1996], dolphins [@Smolker1997], and octopuses [@Finn2009] occurred 
decades after significant advances on other more easily measurable aspects of 
their biology. Since then, scholars have proposed a clear operational 
definition of tool use applicable to all species --- the manipulation of an 
unattached object as an extension of the animal's body to achieve a 
goal [@Shumaker2011] --- and have used the distribution of species meeting this 
definition to make various claims about the evolutionary drivers of tool use 
behaviours. For example, based on the observation that bird species with 
reported tool use tend to have larger brains, researchers have identified 
higher relative brain size as a likely precondition for tool using 
capabilities [@Audet2024;@Lefebvre2002;@Lefebvre1997;@Lefebvre2004]. These 
researchers argue that larger brains are better able to integrate visual and 
somatosensory information when innovating novel behaviours, such as tool use, 
in changing environments [@Sol2005;@Sol2009] (but see ref [@Kaplan2020]). 
Similarly, researchers have used existing reports of tool use in birds to debate
the roles of generalist versus specialist feeding strategies in driving the 
evolution of tool use, with some arguing that feeding generalists require 
technical innovations to expand their dietary niche 
[@Ducatez2015;@Overington2011;@Sol2005] and others arguing that feeding 
specialists require technical innovations for extractive foraging of specific 
foods [@HenkeVonDerMalsburg2020;@MettkeHofmann2014].

However, before we can make claims of this kind, we need to know whether current 
research effort in the literature is sufficient for robust conclusions to be 
drawn about the evolution of tool use. In fact, evidence suggests that research 
effort is often systematically biased towards particular taxonomic groups, parts 
of the world that are easy to access, and species with life history traits that 
make them easier to study, such as larger distribution ranges and population 
sizes [@Ducatez2014]. This is a crucial limitation because insufficient 
observation may lead researchers to miss true instances of tool behaviours and 
thus draw premature conclusions about the evolutionary drivers and origins of 
tool use. Researchers are keenly aware of this problem and have attempted to 
deal with it in different ways, such as controlling for the number of scientific 
papers published on different species. But previous work has not yet attempted 
to quantify and explicitly model the relationship between actual tool-using 
behaviour and what is reported in the scientific literature. If more tool-using 
species exist than previously thought, this could have important implications 
for theories of the evolutionary drivers and origins of tool use and for our 
understanding of how rare this behaviour actually is.

@@

```{r echo=F}
tar_load(crowdData)
trueCrowdData <- dplyr::filter(crowdData, ToolUse == "Y")
bordCrowdData <- dplyr::filter(crowdData, ToolUse == "B")
```

@@

Of the `r trueCrowdData %>% filter(Species != "Unknown") %>% pull(Species) %>% unique() %>% length()` 
parrot species performing true tool use, 
`r trueCrowdData %>% filter(Species != "Unknown") %>% group_by(Species, IndividualID) %>% summarise(n = n(), .groups = "drop") %>% group_by(Species) %>% summarise(check = n() > 1, .groups = "drop") %>% pull(check) %>% sum()` 
were represented in our video survey by two or more individuals over multiple 
independent observations. True tool use always involved the subject using an 
object for self-scratching (`r trueCrowdData %>% mutate(BodyPartScratched = str_detect(BodyPartScratched, "Head/Neck")) %>% pull(BodyPartScratched) %>% sum()` 
videos involved scratching the head and/or neck). The most common tool (`r sum(pull(trueCrowdData, EvaluatedTool) == "Feather")`
videos) was a moulted feather. Human-made objects (e.g., pens, spoons, pieces of 
wood, cardboard) were also common. In `r sum(readxl::read_xlsx(tar_read(crowdsourcing_file))$Directionality == "Tool to body", na.rm = TRUE)` 
videos, parrots manipulated the tool while keeping their body still, rather than 
holding the tool still and moving towards it.

@@

In our model, we infer the tool-using status of each species by allowing each 
species to have their own probability of being a non-tool-user. Following our 
causal model (Figure \@ref(fig:plotDAG)), we predict these probabilities based 
on feeding strategy, encephalisation quotient, and phylogenetic history (see 
Methods for full model). The model additionally takes research effort into 
account by allowing that, among species for which tool use is unobserved, all 
else being equal those with fewer published papers and fewer video search hits 
have a higher probability of being undetected tool users (Figure 
\@ref(fig:plotSurvCure3)).

@@

We found that this phylogenetic survival cure model was able to adequately 
distinguish between species with and without evidence for tool use, with an 
area-under-the-curve classification statistic of 
`r printnum(tar_read(auc), digits = 2)` (Figure 
\@ref(fig:plotSurvCure9)). To further estimate the accuracy of the model's 
predictions, we also used a leave-one-species-out approach with known tool 
users. For each of the 25 tool-using species that were represented on the 
phylogeny and for which we had brain size and genomic data (we lacked data for 
three tool-using species), we fitted the model to a modified dataset which set 
tool use to be absent for the target species in both the scientific literature 
and the video survey. Across `r length(tar_read(cv))` cross-validation models, 
`r sum(tar_read(cv))` models (`r printnum(mean(tar_read(cv))*100, digits = 0)`%) 
continued to predict the target species as having a median posterior probability 
of tool use that was within the range of all other tool users. This 
classification rate was greater than the baseline classification rate of 
`r printnum(mean(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274) * 100, digits = 0)`% 
for species without evidence of tool use in the full model
(`r sum(plotSurvCure1$data %>% filter(ToolUse == "Absent") %>% pull(med) > 0.16478274)` 
of `r nrow(plotSurvCure1$data %>% filter(ToolUse == "Absent"))` species without 
evidence of tool use had a median posterior probability of tool use that was 
within the range of the tool-using species). Together, the area-under-the-curve 
statistic and the leave-one-species-out approach suggest that the model is able 
to adequately classify known tool users, with some error.

Figure \@ref(fig:plotSurvCure1) visualises the ranked posterior probabilities of 
tool use from the phylogenetic survival cure model for all parrot species. As 
expected, the known tool users are ranked towards the top of this list. However, 
several "tool use absent" species also rank highly on the list, despite not 
being identified as tool users in the scientific literature or in our video 
survey. In fact, according to the model, the most likely tool user is a species 
for which tool use is unobserved in our data: the blue-eyed cockatoo (_Cacatua 
ophthalmica_). This species is endemic to Papua New Guinea and is relatively 
understudied, with only `r tar_read(d)$EX.Pre.Tool.Papers[165]` published papers 
and `r tar_read(d)$SearchHits[165]` video search hits, which is fewer than the 
model expects are necessary to discover tool use when it is present (Figure 
\@ref(fig:plotSurvCure2)). This species is also found in the _Cacatua_ genus, a 
clade containing several known tool users. This prediction makes sense given the 
high phylogenetic signal for tool use reported by the model (
Figures \@ref(fig:plotSurvCure5) and \@ref(fig:plotSurvCure6)). Beyond the 
blue-eyed cockatoo, other highly ranked species without observed evidence of 
tool use are the Meyer's parrot (_Poicephalus meyeri_), the golden parakeet 
(_Guaruba guarouba_), the long-billed corella (_Cacatua tenuirostris_), the 
Solomons cockatoo (_Cacatua ducorpsii_), the red-fronted parrot (_Poicephalus 
gulielmi_), the Cape parrot (_Poicephalus robustus_), the yellow-eared parrot 
(_Ognorhynchus icterotis_), the red-vented cockatoo (_Cacatua haematuropygia_), 
and the gang-gang cockatoo (_Callocephalon fimbriatum_). Figure 
\@ref(fig:plotPhylo1) plots these species on the parrot phylogeny, using the top 
ten highest ranked species without observed evidence of tool use as an arbitrary 
cutoff for visualisation purposes.

@@

The predicted probabilities from our phylogenetic survival cure model have 
implications for inferences about the evolutionary drivers and origins of tool 
use in the parrot order. Regarding the drivers of tool use hypothesised in 
Figure \@ref(fig:plotDAG), the phylogenetic survival cure model revealed that 
encephalisation quotient strongly positively predicted the probability of tool 
use (median posterior log odds slope = `r printnum(median(-postFull$bEQ))`, 95% 
CI [`r printnum(quantile(-postFull$bEQ, 0.025))` 
`r printnum(quantile(-postFull$bEQ, 0.975))`]; total _n_ = `r nrow(tar_read(d))` 
species; Figure \@ref(fig:plotSurvCure4)). This helps explain the ranking in 
Figure \@ref(fig:plotSurvCure1): the blue-eyed cockatoo has the largest relative 
brain size in the dataset. We also found that feeding generalist species were 
slightly more likely to be tool users, though the posterior difference between 
generalists and specialists was quite uncertain (median posterior log odds 
difference = `r printnum(median(diff))`, 95% CI [
`r printnum(quantile(diff, 0.025))` `r printnum(quantile(diff, 0.975))`]; total 
_n_ = `r nrow(tar_read(d))` species). These results from the survival cure model 
differed from the results of models fitted to pre-video-survey and 
post-video-survey data without the survival cure component, which found no 
effect of relative brain size and no difference between feeding strategies, 
respectively (Figure \@ref(fig:plotComparison)).

@@

Regarding the origins of tool use, we fitted exploratory ancestral state 
reconstruction models to the pre-video-survey data, the post-video-survey data, 
and the predicted probabilities from the phylogenetic survival cure model. The 
discoveries from our video survey and from our phylogenetic modelling increased 
the likelihood that the capacity for tool use was present in the most recent 
common ancestors for several parrot genera. These include the most recent common 
ancestor of amazon parrots native to the Americas (_Amazona_), the most recent 
common ancestor of the true white cockatoos and corellas found in South East 
Asia and Australasia (_Cacatua_), the most recent common ancestor of the kea and 
the kākā from New Zealand (_Nestor_), and the most recent common ancestor for
the _Poicephalus_ genus native to Africa (Table \@ref(tab:tableASR); 
Figures \@ref(fig:plotASR1) - \@ref(fig:plotASR3)).

@@

These findings have a number of important implications. First, our findings show 
that current research effort in the scientific literature is insufficient to 
capture the real world occurrence of parrot tool use. If the scientific 
literature had sampled the natural world sufficiently, we would expect to see 
close correspondence between those species reported as tool users in the 
literature and those species the public have filmed performing tool use. 
Instead, we discovered a large discrepancy between these two data sources, both 
in the prevalence of tool use and the species identified, likely due to the 
numerous difficulties inherent to conducting observations in the wild. This 
raises the possibility that other rare behaviours remain yet to be discovered in 
the scientific literature.

Second, in terms of the evolution of tool use in parrots, our study challenges a 
key assumption made in the literature to date: that only a minority of parrots 
are tool users [@Auersperg2012;@Bastos2021;@Goodman2018;@Lambert2015;@OHara2021]. 
The paucity of evidence for tool use across parrots in the literature initially 
implied that the capacity for tool use may have evolved independently in 
different parrot species. Our discovery of the widespread distribution of tool 
use across the parrot phylogeny, along with the strong phylogenetic signal in 
this expanded dataset, challenges this and suggests that, at least for some 
parrot clades, the capacity for tool use might be a homologous trait that has 
been evolutionarily conserved. Our exploratory ancestral state reconstruction 
analysis provides preliminary support for this hypothesis, revealing an 
increased probability of the capacity for tool use among the most recent common 
ancestors for the _Amazona_, _Cacatua_, _Nestor_, and _Poicephalus_ genera. The 
complexity of tool use behaviour that such a basic capacity might afford, and 
the conditions under which it emerges in the wild, are yet to be established. 
However, even at this preliminary stage, our analysis raises an alternative 
hypothesis for the observed tool use in _Cacatua_ [@Auersperg2012;@OHara2021] 
and _Nestor_ [@Auersperg2009;@Auersperg2010;@Auersperg2011a;@Auersperg2011b;
@Bastos2021;@Goodman2018], namely that tool behaviours have arisen due to the 
common ancestor having the capacity to use tools, rather than from independent 
evolution or behavioural innovation within species.

@@

All instances of tool use in our video survey met the established criteria for 
tool use in the literature: the manipulation of an unattached object as an 
extension of the animal's body to achieve a goal [@Shumaker2011]. However, a 
striking feature of the dataset is that the 17 additional species identified in
the video survey were exclusively observed using tools for self-scratching.
While parrots in the wild use tools to achieve a variety of goals, including 
extractive foraging and preening [@Goodman2018;@OHara2021], parrots in captivity
only require tools to achieve the latter goal. It could be argued that 
self-scratching is a more cognitively simplistic form of tool use than 
extractive foraging as it is egocentric, meaning that the tool is directed 
towards oneself rather than an external object. Nevertheless, self-scratching 
meets the definition for a more complex type of embodied tool use known as 
'tooling', the deliberate generation of a mechanical interface by using an 
object to manipulate another target or surface [@Fragaszy2018]. In our dataset, 
most self-scratching tooling instances involved manipulating the tool while 
keeping the body still, suggesting that these were goal-directed, deliberate 
movements by individuals. This makes it difficult to dismiss self-scratching as 
unimportant: to do so would require a redefinition of tool-use and tooling in 
non-human animals. Moreover, the fact that we found strong phylogenetic 
clustering of self-scratching tool use from the video survey with other examples 
of tool use from the literature supports a common underlying cognitive 
mechanism. In line with this, some of the species of parrots that use tools for 
self-scratching in captivity also use tools for other purposes in the wild 
[@Auersperg2012;@Bastos2021;@Goodman2018;@OHara2021].

A potential concern with data originating from digital video platforms is its 
reliability [@Auersperg2020;@DechaumeMoncharmont2020;@Farrar2020;@Nelson2013;@Sandor2020;@vonBayern2020]. 
In particular, the self-care tool use instances detected in our video survey 
might be attributable to training or unintentional cueing by the birds' owners.
However, there was little evidence to suggest that the observations of 
self-scratching tool use in the video survey were merely unintentional accidents 
or explicitly trained behaviours. Individual parrots often used tools slowly and 
repetitively over long periods of time, even across multiple different videos, 
suggesting that their behaviour was not random or accidental [@Sandor2020]. 
Parrots preferentially employed self-scratching tools on areas of their body 
that were otherwise inaccessible, with 96% of all instances involving scratching 
of the head or neck, suggesting intentional tool use. For 60% of the species in 
the video survey, we found two or more videos of repetitive and sustained 
scratching by different individuals of the same species in separate households, 
suggesting that the manipulations were intentional and recurring events that did 
not represent unusual stereotypies of any single individual. Regarding the 
possibility of training or cueing from owners, over half of the videos contained 
no evidence of human interaction aside from filming the behaviour. Humans only 
handed parrots their tools in two of the videos, and none of the videos featured 
owners directly rewarding tool use behaviours with food. It is unlikely that 
owners would specifically train parrots to scratch themselves with tools. 
Finally, the high levels of phylogenetic signal in our data provide strong 
evidence that the observations from our video survey reflect
biologically-endowed capacities for tool use rather than accidental or trained
behaviours, which would likely appear uniformly across the phylogeny. 
Nevertheless, future work using data such as this should choose target 
behaviours carefully and examine videos for evidence of training or cueing.

Our data reveals that a wide range of parrot species exhibit self-scratching 
tool use behaviours in captive environments. It is possible that these self-care 
behaviours emerge more often in captivity because pet parrots are kept with few 
or no conspecifics, and therefore must innovate tool use to preen inaccessible 
body areas, such as their heads and necks, which would have been otherwise 
preened by flock members. This may indicate that human interaction and 
allopreening in the absence of a conspecific is insufficient for pet parrots, as 
suggested by studies on welfare and chronic stress in lone captive parrots 
[@Aydinonat2014;@Engebretson2006;@Mason2010;@Meehan2003]. Beyond social 
isolation in captivity, other factors that reduce preening efficacy by an 
individual might also lead to the innovation of self-care tooling: for example,
a captive kea housed with conspecifics in a naturalistic environment was shown
to have innovated the use of a pebble tool for self-care likely as a consequence 
of a missing upper mandible [@Bastos2021]. These behaviours might therefore be 
innovated by individuals when ecologically necessary. Critically, however, a 
cognitive capacity to innovate a behaviour such as tool use must exist in a 
species in order to manifest in any environment, whether that be captive or in 
the wild. As such, despite differences in individuals' life experiences in 
captive and wild settings, such behaviour emerging in a captive setting at the 
very least indicates an underlying cognitive capacity to innovate that same 
behaviour in the wild.

@@
